---
title: "Statistics 650 - Project"
author: "Nurahmet Emin"
date: "Wednesday October 13, 2019"
output:
  word_document: default
  html_notebook: default
---
```{r}

library(tidyverse)
library(ggmap)
library(skimr)
library(lubridate)
library(textreadr)
library(nycflights13)
```

##Question1.

**Answer:**

I have used the data of months Jan-May 2018 as below.

```{r message=FALSE}
flight_bay201801 <- read_csv(file="./data/On_Time_Reporting_Carrier_On_Time_Performance_(1987_present)_2018_1.csv")
flight_bay201802 <- read_csv(file="./data/On_Time_Reporting_Carrier_On_Time_Performance_(1987_present)_2018_2.csv")
flight_bay201803 <- read_csv(file="./data/On_Time_Reporting_Carrier_On_Time_Performance_(1987_present)_2018_3.csv")
flight_bay201804 <- read_csv(file="./data/On_Time_Reporting_Carrier_On_Time_Performance_(1987_present)_2018_4.csv")
flight_bay201805 <- read_csv(file="./data/On_Time_Reporting_Carrier_On_Time_Performance_(1987_present)_2018_5.csv")

```


```{r}
bay_flight01<- flight_bay201801%>% filter(Origin=="SFO" | Origin=="SJC"| Origin=="OAK")%>%select(c(Year:Reporting_Airline,Tail_Number,Origin:OriginState,Dest:DestState,CRSDepTime:DepDel15,CRSArrTime:ArrDel15,AirTime,Distance))
bay_flight02<- flight_bay201802%>% filter(Origin=="SFO" |Origin=="SJC"|Origin=="OAK")%>%select(c(Year:Reporting_Airline,Tail_Number,Origin:OriginState,Dest:DestState,CRSDepTime:DepDel15,CRSArrTime:ArrDel15,AirTime,Distance))
bay_flight03<- flight_bay201803%>% filter(Origin=="SFO" | Origin=="SJC"|Origin=="OAK")%>%select(c(Year:Reporting_Airline,Tail_Number,Origin:OriginState,Dest:DestState,CRSDepTime:DepDel15,CRSArrTime:ArrDel15,AirTime,Distance))
bay_flight04<- flight_bay201804%>% filter(Origin=="SFO" | Origin=="SJC"|Origin=="OAK")%>%select(c(Year:Reporting_Airline,Tail_Number,Origin:OriginState,Dest:DestState,CRSDepTime:DepDel15,CRSArrTime:ArrDel15,AirTime,Distance))
bay_flight05<- flight_bay201805%>% filter(Origin=="SFO" | Origin=="SJC"|Origin=="OAK")%>%select(c(Year:Reporting_Airline,Tail_Number,Origin:OriginState,Dest:DestState,CRSDepTime:DepDel15,CRSArrTime:ArrDel15,AirTime,Distance))

```

##Question2.

**Answer:**

There were 22606 flights in Jan 2018 flewed from the three airports in bay area.
```{r}
dim(bay_flight01)
```


```{r}
sfoflights18<- bind_rows(bay_flight01,bay_flight02,bay_flight03,bay_flight04,bay_flight05)
dim(sfoflights18)
head(sfoflights18)


```

##Question3.

**Answer:**

```{r}
nycflights13_names <- tribble(
  ~columns, ~description,
  'year', 'Date of departure. - year',
  'month', 'Date of departure. - month',
  'day', 'Date of departure.',
  'dep_time', 'Actual departure and arrival times (format HHMM or HMM), local tz.',
  'arr_time', 'Actual departure and arrival times (format HHMM or HMM), local tz.',
  'sched_dep_time', 'Scheduled departure and arrival times (format HHMM or HMM), local tz.',
  'sched_arr_time', 'Scheduled departure and arrival times (format HHMM or HMM), local tz.',
  'dep_delay', 'Departure and arrival delays, in minutes. Negative times represent early departures/arrivals.',
  'arr_delay', 'Departure and arrival delays, in minutes. Negative times represent early departures/arrivals.',
  'carrier', 'Two letter carrier abbreviation. See airlines to get name.',
  'flight', 'Flight number.',
  'tailnum', 'Plane tail number. See planes for additional metadata.',
  'origin', 'Origin and destination. See airports for additional metadata.',
  'dest', 'Origin and destination. See airports for additional metadata.',
  'air_time', 'Amount of time spent in the air, in minutes.',
  'distance', 'Distance between airports, in miles.',
  'hour', 'Time of scheduled departure broken into hour and minutes. - hour',
  'minute', 'Time of scheduled departure broken into hour and minutes. minute',
  'time_hour', 'Scheduled date and hour of the flight as a POSIXct date. Along with origin, can be used to join flights data to weather data.'
)

kable(nycflights13_names)
```




```{r}
X <- read_html("readme.html")

head(X)

# Remove the top four lines.

X <- X[-c(1,2,3,4,5)]

head(X)

# Now create a matrix that contains the variable names in the first column and the description in the second column.

Y <- matrix(X[1:218], ncol = 2, byrow = TRUE)

head(Y)

on_time_names <- tibble(columns = Y[,1], description = Y[,1])
on_time_names
 
on_time_names <- on_time_names %>% mutate(columns = tolower(columns))
head(on_time_names)

```

##Question4.

**Answer:**

Below is the table with the new variables available in flightsfo18 dataset.
```{r}

sfoflights18_names <- tribble(
  ~columns, ~description,
  'Quarter', 'Quarter (1-4)',
  'DayOfWeek', 'Day of Week',
  'OriginCityName', 'Origin Airport, City Name',
  'OriginStat', 'Origin Airport, State Code',
  'DepDelayMinutes','Difference in minutes between scheduled and actual departure time. Early departures set to 0.',
  'DestCityName', 'Destination Airport, City Name',
  'DestState', 'estination Airport, State Code.',
  'DepDel15', 'Departure Delay Indicator, 15 Minutes or More (1=Yes).',
  'ArrDelayMinutes', 'Difference in minutes between scheduled and actual arrival time. Early arrivals set to 0.',
  'ArrDel15', 'Arrival Delay Indicator, 15 Minutes or More (1=Yes'
  )

kable(sfoflights18_names)
```

##Question5(4.2).

**Answer:**

In the data for Jan-May 2018, March had the highes cancelled flights. Feb had the lowest cancelled flights. Maybe because of the rainy weather in 2018 from my memory.
```{r}
sfoflights18%>%select(Month,ArrDelay)%>% group_by(Month)%>%skim()
```

##Question5(4.3).

**Answer:**

Flight N633VA travelled the most.

```{r}
sfoflights18 %>%select(Tail_Number) %>%group_by(Tail_Number)%>%count(count=n()) %>%arrange(desc(n))
```
```{r}
sfoflights18 %>%select(Year, Month, Tail_Number) %>%filter(Tail_Number=="N633VA") %>%group_by(Year,Month) %>%summarise(count=n())
```

```{r}

sfoflights18 %>%filter(Tail_Number=="N633VA") %>%group_by(Month)%>%tally()%>%ggplot(aes(x= Month, y=n))+geom_point()+geom_line()
```

Number of Trips per Week 

```{r}
sfoflights1 <- sfoflights18 %>% mutate( ymd = as.Date(paste(Year, Month, DayofMonth, sep='-'))) %>%
  mutate( week = as.integer(format(as.Date(ymd), "%U"))  ) %>% 
  filter(Tail_Number=="N633VA") %>%
  group_by(week) %>%
  tally()
sfoflights1
sfoflights1%>% ggplot(aes(x=week, y=n))+geom_point()+geom_line()+geom_smooth()
```

##Question5(4.4).

**Answer:**
3,322 planes flew from SFO are included in planes table.

```{r}
planes_1<- planes%>%select(tailnum,year)%>%group_by(year)%>%arrange((year))
planes_1
planes2 <-planes %>%rename( year_m =year )
planes2%>% left_join(sfoflights18, by= c("tailnum"="Tail_Number"))%>%group_by(tailnum)%>% count(n=n())
```



